My First R Markdown
===================
###Loading and preprocessing the data
Unzip and read the file
```{r}
#Turn off warnings to clean up output
options(warn = -1)
data <- read.csv(unz("activity.zip", "activity.csv"))
```
Convert factor to date using lubridate
```{r}
library(lubridate)
data$date <- ymd(data$date)
```
###What is mean total number of steps taken per day?
Summarize data using dplyr
```{r}
#Turn off messages to clean up output
suppressMessages(library(dplyr))
a <- summarize(group_by(data, date), sum(steps))
names(a) <- c("date", "steps")
```
Create histogram using ggplot2
```{r}
library(ggplot2)
ggplot(a, aes(steps)) + geom_histogram(binwidth = 1000) 
```

Calculate the mean and median
```{r}
mean(a$steps, na.rm = TRUE)
median(a$steps, na.rm = TRUE)
```
###What is the average daily activity pattern?
Summarize data using dplyr
```{r}
b <- summarize(group_by(data, interval), mean(steps, na.rm = TRUE))
names(b) <- c("interval", "steps")
```
Create time series using ggplot2
```{r}
ggplot(b, aes(interval, steps)) + geom_line()
```

Find interval with maximum steps
```{r}
b[which(b$steps == max(b$steps)),1]
```
###Imputing missing values
Calculate number of missing values
```{r}
sum(!complete.cases(data))
```

```{r}
#Turn warnings back on just in case
options(warn = 0)