---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Unzip and read the file
```{r}
#Turn off warnings to clean up output
options(warn = -1)
data <- read.csv(unz("activity.zip", "activity.csv"))
```
Convert factor to date using lubridate
```{r}
library(lubridate)
data$date <- ymd(data$date)
```

## What is mean total number of steps taken per day?
Summarize data using dplyr
```{r}
#Turn off messages to clean up output
suppressMessages(library(dplyr))
a <- summarize(group_by(data, date), sum(steps))
names(a) <- c("date", "steps")
```
Create histogram using ggplot2
```{r}
library(ggplot2)
ggplot(a, aes(steps)) + geom_histogram(binwidth = 1000) 
```

Calculate the mean and median
```{r}
mean(a$steps, na.rm = TRUE)
median(a$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?
Summarize data using dplyr
```{r}
b <- summarize(group_by(data, interval), mean(steps, na.rm = TRUE))
names(b) <- c("interval", "steps")
```
Create time series using ggplot2
```{r}
ggplot(b, aes(interval, steps)) + geom_line()
```

Find interval with maximum steps
```{r}
b[which(b$steps == max(b$steps)),1]
```

## Imputing missing values
Calculate number of missing values
```{r}
sum(!complete.cases(data))
```
Imput missing (NA) values using the mean of each interval. We have already calculated those means in b. Use b to create a new dataset called x.
```{r}
#Use merge to match mean values with intervals
x <- merge(data, b, by = "interval")
#Then use arrange to restore row sort by date then interval
x <- arrange(x, date, interval)
#Then use mutate to add a column which chooses the mean if there's an NA else the original measurement
x <- mutate(x, steps = ifelse(is.na(steps.x), steps.y, steps.x))
#Then use select to drop unnecessary columns and restore column order
x <- select(x, steps, date, interval)
```
The new dataset has no incomplete cases.
```{r}
sum(!complete.cases(x))
```
Summarize data and create histogram
```{r}
a2 <- summarize(group_by(x, date), sum(steps))
names(a2) <- c("date", "steps")
ggplot(a2, aes(steps)) + geom_histogram(binwidth = 1000) 
```

Compare density plots between original dataset (with NA's) and imputed dataset. This is the impact of imputing data--at least via this method of imputation.
```{r}
#create combined summary dataset in order to place two density plots in one plot
a$status <- 'original'
a2$status <- 'imputed'
a3 <- rbind(a,a2)
ggplot(a3, aes(steps, fill = status)) + geom_density(alpha = 0.3)
```

Calculate means and medians for the two datasets for comparison purposes.
```{r}
mean(a$steps, na.rm = TRUE)
mean(a2$steps, na.rm = TRUE)
median(a$steps, na.rm = TRUE)
median(a2$steps, na.rm = TRUE)
```
## Are there differences in activity patterns between weekdays and weekends?
Add weekday/weekend factor variable to imputed dataset and create a new dataset for simplicity
```{r}
y <- mutate(x, weekday = weekdays(date, abbreviate = TRUE))
y <- mutate(y, weekpart = ifelse((weekday == "Sat" | weekday == "Sun"), "weekend", "weekday"))
y$weekpart = as.factor(y$weekpart)
#Relevel to replicate panel plot order
y$weekpart = relevel(y$weekpart, "weekend")
```
Summarize data and create time series plot
```{r}
z <- summarize(group_by(y, weekpart, interval), mean(steps))
names(z) <- c("weekpart", "interval", "steps")
ggplot(z, aes(x = interval, y = steps)) +geom_line() + facet_wrap(~weekpart, nrow = 2)
```{r}
#Turn warnings back on just in case
options(warn = 0)